Toward Supervised Reinforcement Learning with Partial States for Social HRI

### 2017-08-22 paper.tex
Fix typos and paragraph normalisation

\maketitle
\begin{abstract}
%As robots will integrate human society, they will have to satify humans'
%expectations. However at the design time, it seems hardly possible to
%know in advance exactly all the behaviours expected from the robot. As such to
%be accepted in human society robots have to learn how to interact with humans.
    Social interacting is a complex task for which machine learning holds particular promise. However, as no sufficiently
    accurate simulator of human interactions exists today, the learning of social interaction strategies has to happen online in the real
    world. Actions executed by the robot impact on humans, and as such
    have to be carefully selected, making it impossible to rely on random exploration. Additionally, no clear reward function exists for
    social interactions. This implies that classical approaches such as
    Reinforcement Learning cannot be directly applied for learning how to
    interact with the social world. As such we argue that
    robots will profit from human expertise and guidance to learn social
    interactions. And as the quantity of input a human can provide is limited,
    new methods have to be designed to use human input more efficiently. In this paper we describe a setup in which We
    combine a framework called Supervised Progressively Autonomous Robot Competencies (SPARC), which
    allows safer online learning with Reinforcement Learning, with the use
    partial states rather than full states to accelerate generalisation and
    obtain more quickly a usable action policy.
\end{abstract}
%===============================================================================

@@

Human-Robot Interaction (HRI) studies how people and robot can co-exist in society, and how they can interact socially in different environments and
contexts. Robot are expected to behave appropriately regardless of the domain
of interaction. However, it is impossible to foresee all possible interaction outcomes in dynamic and open social domains, as such the robot's responses cannot be implemented in the robot before its deployment in the real world. Similar to
how people learn, robots need to be able to learn how to complete tasks through learning and optimising action policies. This including learning social norms and how to make sense of the social world.

@@

\begin{comment}
This approach in especially relevant for HRI as gathering datapoints has to be
performed in the real world and interacting with humans forces a robot to have
an appropriate action policy at anytime of the interaction. even in the early
phase when the robot has no information about the way its environment react to
its actions.

When interacting in the real world, humans face huge spaces of almost unlimited
dimensions, and despite this complexity, humans, and especially children are able
to learn quickly, often from only a few demonstrations or examples. One of the
 reason could be that they are able to take the most out of every datapoint available to them. 
 On the other
hand, recent impressive advances in Machine Learning such as Deep Learning~\cite{lecun2015deep}
have been made possible by the availability of huge dataset covering millions
of labeled examples or dozens of hours of recording. These progress have
targetting either perception, where labeled datapoints can be available or
action selection in virtual environments where the impact of actions is
limited.

However, these large labelled dataset or virtul worlds do not exist for
social interaction. To learn how to interact with humans, robots have to
experiment in the real world facing the actual reactions from their actions. As
such they
have to learn quickly while consistently exhibiting a acceptable behaviours.
One way of ensuring this correct  behaviours is to rely on a human to supervise
the robot during the early phase of learning, and this human can also provide
inputs to boostrap the learning, achieving a faster learning. This paper present
a novel method based on human supervision, feature highlighting and partial
state reinforcement learning to achieve this fast and safe learning.
\end{comment}
%===============================================================================

@@

%Classic approaches in RL generally rely on exploring virtual environments where the agent
%tries to maximise the reward.
In many cases where RL achieves success, the agent
has access to a virtual environment where the only real cost of exploring is 
computational effort: the agent can interact as long as needed to gather
enough information on the environment and the result of its actions on the environment in order to find a sufficiently optimal action policy. Using simulation RL can achieve impressive results, as shown in RL learning to play Backgammon in the
90s~\cite{tesauro1995temporal} to the more recent success in mastering the game
of Go~\cite{silver2016mastering}. Even when a virtual environment is available,
but especially when it is not, human knowledge and effort is required during the
design of the algorithm or the learning phase before the learning can be successful.

@@

SPARC gives control of the robot's actions to an expert, who can guide the
exploration in the desired part of the environment which ensures the robot's
behaviour is consistently appropriate. As the exploration is
guided, and all the actions are useful, the learning can be faster than
autonomous learning or feedback based learning as illustrated in Figure
\ref{fig:comparison}.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{./figs/motivation.pdf}
    \caption{An illustration of the evolution over time of the performance, autonomy and human workload for a
    an autonomous learner, an approach using human rewards, and SPARC.}
    \label{fig:comparison}
\end{figure}

@@

In \cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot, making the assumption that every action has been explicitely or implicitely approved by the supervisor. However this method
was  directly mapping a single (state, action) pair to a reward without making
use of any kind of generalisation. As such it was not applicable to
environments with a continuous or high dimensional space or non-deterministic
transitions from one state to another, elements which are typically present in social interactions.

\section{Partial State Supervised Reinforcement Learning}
\begin{comment}
Humans evolve in a highly complex environment with a nearly infinite number of
dimensions originating from different modality through our senses. But, despite
this complexity, we can select actions according to important features of the
state, regardless of other parts of it.

Taking inspiration from this observation, we argue that focusing on partial states, rather
than full states can both increase the speed of learning for an agent and
provide some of the generalisation required to interact in complex environment.
Guidance from a human can highlight important features in the state space and
associate them to actions. This shifts the state action pair paradigm to a
partial state action pair, and in the case of RL, the tuple: partial state,
action and reward. 
\end{comment}

@@

\begin{comment}
%===============================================================================
\section{Methodology}

Similarly to a SPARC setup, the supervisor can select actions, or when presented
with an action about to be executed by the robot either cancel it or let
it be executed.

In addition to these actions, the supervisor can use the GUI to select features
of the environment that can be sent to the algorithm to reduce the state associated
to this action to a partial state. Similarly, the algorithm can indicate which
parts of the states have been used for the selection to the supervisor who can
correct them while the action is being executed.

%Example?
\end{comment}

@@

To each action
$a \in \mathcal{A}$ we can associate a set $\mathcal{C}_{a}$ of pairs $(s',r)$
representing the rating done by the supervisor to action $a$ with features
highlighted for the multiple $s'$. When adding a new pairt $(s',r)$, we can discard potential
previous pairs with an identical $s'$ to represent the evolution of the policy
evaluation by the supervisor.

@@

An example of an application is a social robot which interacts with children in an educational scenario.
The robot plays an educational game with children to teach them notions about
diverse topics according to the needs of the teacher. For this example, a child
and a robot are playing a game about the food web, teaching wich animals eat which
ones on a Sandtray~\cite{baxter2012touchscreen}. In addition to the child and
the robot, an adult supervises the robot using a tablet with a Graphical User
Interface (GUI) to teach the robot how to interact with the child as shown in
Figure~\ref{fig:setup}.

@@

%===============================================================================
\begin{comment}
\section{Challenges}

A large body of the
Interactive Machine Learning research is focused on simple inputs from the
humans, such as providing labels \cite{cakmak2012designing}, rewards
\cite{knox2009interactively} or kinesthetic demonstrations
\cite{billard2008robot}. However, as stated in \cite{amershi2014power}: "people
want to demonstrate how learners should behave" and "users may desire richer
control over machine-learning systems than simply labeling data". Human teachers
want to be more active and involved in the teaching process, to make their
student learning faster. 
And, as shown in \cite{senft2017supervised}, providing teacher with this control
can keep the agent in relevant states of the environment, speeding the learning
and reducing the risk to have the robot execute undesired actions.

In human social learning, both interactants have access to a similar set of
inputs: the senses available to a human body, and the teacher has additional
knowledge about the environment that can be communicated in an open ended way to
the learner. On the other in human-robot teaching, the way to sense the world
can be highly different between the two partners, and as of today, there is no
open ended way of communicating new concepts or explication as natural langage
processing is not yet solve. This creates two main challenges to make robot
learning for human use complex inputs: 
\begin{itemize}
    \item Having a mapping between the action state space of the human and the
            robot ensuring that feature used for action selection for one
            interactant are available to the other.
    \item Communicating these features from one participant to the other: allow
        the robot to make sense of the human's instructions and the human to
        understand the robot's intentions.
\end{itemize}
\end{comment}
\section{Future work}

@@

The current work could also be extended to allow the agent to continue to improve its
behaviour even in the absence of a supervisor, progressively exploring around the
learnt policy improving its behaviour beyond the demonstrations. This
could be done by allowing the supervisor to provide rewards during the learning
phase and combine these rewards with the demonstrations to learn a reward
function in a fashion similar to Inverse Reinforcement Learning
\cite{abbeel2004apprenticeship} or TAMER \cite{knox2009interactively}. This
could also use partial states associated to these rewards to ease the
generalisation of the reward function with only a low number of datapoints.
But without the supervisor, the assumption that only a
myopic action selection is sufficient would not hold anymore and the problem of
delayed rewards would have to be tackled.
%===============================================================================
%\section{Conclusion}
%\label{sec:conclusion}
%==============================================================================
\section{Acknowledgments}
This work was supported by the EU FP7 DREAM project (grant no.  611391) and EU
H2020 Marie Sklodowska-Curie Actions project DoRoThy (grant 657227).

>>>>

\maketitle \begin{abstract}
%As robots will integrate human society, they will have to satify humans'
    %expectations. However at the design time, it seems hardly possible to know
    %in advance exactly all the behaviours expected from the robot. As such to
    %be accepted in human society robots have to learn how to interact with
    %humans.
    Social interacting is a complex task for which machine learning holds
    particular promise. However, as no sufficiently accurate simulator of human
    interactions exists today, the learning of social interaction strategies has
    to happen online in the real world. Actions executed by the robot impact on
    humans, and as such have to be carefully selected, making it impossible to
    rely on random exploration. Additionally, no clear reward function exists
    for social interactions. This implies that traditional approaches used for
    Reinforcement Learning cannot be directly applied for learning how to
    interact with the social world. As such we argue that robots will profit
    from human expertise and guidance to learn social interactions. And as the
    quantity of input a human can provide is limited, new methods have to be
    designed to use human input more efficiently. In this paper we describe a
    setup in which we combine a framework called Supervised Progressively
    Autonomous Robot Competencies (SPARC), which allows safer online learning
    with Reinforcement Learning, with the use partial states rather than full
    states to accelerate generalisation and obtain more quickly a usable action
    policy.  
\end{abstract}
%===============================================================================

@@

Human-Robot Interaction (HRI) studies how people and robot can co-exist in
 society, and how they can interact socially in different environments and
 contexts. Robot are expected to behave appropriately regardless of the domain
 of interaction. However, it is impossible to foresee all possible interaction
 outcomes in dynamic and open social domains, as such the robot's responses
 cannot be implemented in the robot before its deployment in the real world.
 Similarly to  people, robots need to be able to learn how to complete tasks
 throug creating and optimising action policies. This includes learning social
 norms and how to make sense of the social world.

@@

===============================================================================

@@

%Classic approaches in RL generally rely on exploring virtual environments where
%the agent tries to maximise the reward.
In many cases where RL achieves success, the agent has access to a virtual
environment where the only real cost of exploring is computational effort: the
agent can interact as long as needed to gather enough information on the
environment and the result of its actions on the environment in order to find a
sufficiently optimal action policy. Using simulation RL can achieve impressive
results, as shown in RL learning to play Backgammon in the
90s~\cite{tesauro1995temporal} to the more recent success in mastering the game
of Go~\cite{silver2016mastering}. Even when a virtual environment is available,
but especially when it is not, human knowledge and effort is required during the
design of the algorithm or the learning phase before the learning can be
successful.

@@

SPARC gives control of the robot's actions to an expert, who can guide the
exploration in the desired part of the environment which ensures the robot's
behaviour is consistently appropriate. As the exploration is guided, and all the
actions are useful, the learning can be faster than autonomous learning or
feedback based learning as illustrated in Figure~\ref{fig:comparison}.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{./figs/motivation.pdf}
    \caption{An illustration of the evolution over time of the performance,
    autonomy and human workload for an autonomous learner, an approach using
    human rewards, and SPARC.}
    \label{fig:comparison}
\end{figure}

@@

In \cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot, making the
assumption that every action has been explicitly or implicitely approved by the
supervisor. However this method was  directly mapping a single (state, action)
pair to a reward without making use of any kind of generalisation. As such it
was not applicable to environments with a continuous or high dimensional space
or non-deterministic transitions from one state to another, elements which are
typically present in social interactions.

\section{Partial State Supervised Reinforcement Learning}

@@

@@

To each action $a \in \mathcal{A}$ we can associate a set $\mathcal{C}_{a}$ of
pairs $(s',r)$ representing the rating done by the supervisor to action $a$ with
features highlighted for the multiple $s'$. When adding a new pair $(s',r)$, we
can discard potential previous pairs with an identical $s'$ to represent the
evolution of the policy evaluation by the supervisor.

@@

An example of an application is a social robot which interacts with children in
an educational scenario.  The robot plays an educational game with children to
teach them notions about diverse topics according to the needs of the teacher.
For this example, a child and a robot are playing a game about the food web,
teaching which animals eat which ones on a Sandtray~\cite{baxter2012touchscreen}.
In addition to the child and the robot, an adult supervises the robot using a
tablet with a Graphical User Interface (GUI) to teach the robot how to interact
with the child as shown in Figure~\ref{fig:setup}.

@@

%===============================================================================
\section{Future work}

@@

The current work could also be extended to allow the agent to continue to
    improve its behaviour even in the absence of a supervisor, progressively
    exploring around the learnt policy improving its behaviour beyond the
    demonstrations. This could be done by allowing the supervisor to provide
    rewards during the learning phase and combine these rewards with the
    demonstrations to learn a reward function in a fashion similar to Inverse
    Reinforcement Learning \cite{abbeel2004apprenticeship} or TAMER
    \cite{knox2009interactively}. This could also use partial states associated
    to these rewards to ease the generalisation of the reward function with only
    a low number of datapoints.  But without the supervisor, the assumption that
    only a myopic action selection is sufficient would not hold anymore and the
    problem of delayed rewards would have to be tackled.
%===============================================================================
    %\section{Conclusion} \label{sec:conclusion}
    %==============================================================================
\section{Acknowledgments} This work was supported by the EU FP7 DREAM project
    (grant no.  611391) and EU H2020 Marie Sklodowska-Curie Actions project
    DoRoThy (grant 657227).
